---
title: "Assignment 5"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

### [Final Project Website]{.ul}

This is my submission for Assignment \#5! I have three interactive plots below that make improvements on my midterm. These plots will be making an appearance on my final project website, which is exciting! You can check out my final website [here](https://jaxabercrombie.github.io/finalproject-pm566/), which features the following three plots and descriptions for each.

```{r libraries, include = FALSE}
library(tidyverse)
library(plotly)
library(cowplot)
library(RCurl)
library(scales)
```

```{r dataprepare, echo = FALSE, include = FALSE}
# Read in the CSV files
vaxCA <- read.csv("CA Vaccine Data.csv")
covidCA <- read.csv("CA Deaths and Cases.csv")

# Merge the first two datasets
covidCA <-
  covidCA %>%
  mutate(county = area)

vaxCA <-
  vaxCA %>%
  mutate(date = as.Date(vaxCA$date, "%m/%d/%Y"))

mergedCA <- merge(vaxCA, covidCA, 
                  by = c("date","county"),
                  all.x = TRUE)

mergedCA <-
  mergedCA %>%
  filter(date >= as.Date("2020-12-15")) %>%
  filter(date <= as.Date("2021-11-03")) %>%
  group_by(date, county) %>% 
  summarise(across(c(total_doses,cumulative_total_doses,
                     pfizer_doses,cumulative_pfizer_doses,
                     moderna_doses,cumulative_moderna_doses,
                     jj_doses,cumulative_jj_doses,
                     partially_vaccinated,
                     total_partially_vaccinated,fully_vaccinated,
                     cumulative_fully_vaccinated,at_least_one_dose,
                     cumulative_at_least_one_dose,population,
                     cases,cumulative_cases,deaths,
                     cumulative_deaths,total_tests,
                     cumulative_total_tests,positive_tests,
                     cumulative_positive_tests), mean, .groups = date))

# Create other variables
mergedCA <-
  mergedCA %>%
  mutate(dose_standard = (cumulative_total_doses/population),
         pfizer_perc = ((cumulative_pfizer_doses/cumulative_total_doses)*100),
         moderna_perc = ((cumulative_moderna_doses/cumulative_total_doses)*100),
         jj_perc = ((cumulative_jj_doses/cumulative_total_doses)*100),
         perc_vaccinated = (cumulative_fully_vaccinated/population)*100,
         perc_partial = (cumulative_at_least_one_dose/population)*100)

```

$$\\[.1in]$$

### [Figure 1: Vaccination Over Time]{.ul}

```{r fig1, echo = FALSE, warning = FALSE}
superior <- c("Butte", "Colusa", "Del Norte", "Glenn", "Humboldt", "Lake", "Lassen", "Mendocino", "Modoc", "Nevada", "Plumas", "Shasta", "Sierra", "Siskiyou", "Tehama", "Trinity")

central <- c("Alpine", "Amador", "Calaveras", "El Dorado", "Fresno", "Inyo", "Kings", "Madera", "Mariposa", "Merced", "Mono", "Placer", "Sacramento", "San Joaquin", "Stanislaus", "Sutter", "Yuba", "Tulare", "Tuolumne", "Yolo")

bay <- c("Alameda", "Contra Costa", "Marin", "Monterey", "Napa", "San Benito", "San Francisco", "San Mateo", "Santa Clara", "Santa Cruz", "Solano", "Sonoma")

southern <- c("Imperial", "Kern", "Orange", "Riverside", "San Bernardino", "San Diego", "San Luis Obispo", "Santa Barbara", "Ventura", "Los Angeles")


mergedCA <- 
  mergedCA %>% 
  mutate(region = case_when(county %in% superior ~ "Superior California",
                            county %in% central ~ "Central California",
                            county %in% bay ~ "Bay Area",
                            county %in% southern ~ "Southern California")) %>%
  mutate(County = county)


plot <- 
    mergedCA %>%
    ggplot() + 
    geom_line(aes(x = date, 
                  y = cumulative_fully_vaccinated/population*100, 
                  color = County)) +
    labs(title = "Percent Fully Vaccinated Since Roll-out by Region", x = "Date", 
         y = "Percent Fully Vaccinated") +
  facet_wrap(~region)

plotly <- ggplotly(plot)

plotly
```

$$\\[.1in]$$

### [Fig 2: Doses, Cases, and Deaths]{.ul}

```{r fig2, echo = FALSE, warning = FALSE}
fig2 <- 
  mergedCA %>%
  ggplot(aes(group = region)) +
  geom_line(mapping = aes(x = date, y = total_doses, color = "Doses Administered")) +
  geom_line(mapping = aes(x = date, y = cases, color = "Case Count")) +
  geom_line(mapping = aes(x = date, y = deaths, color = "Death Count")) +
  labs(title = "Doses, Cases, and Deaths by Region", x = "Date", y = "Count") +
  scale_colour_manual(values = c("Doses Administered"="#C78888", 
                                   "Case Count"="#88C7C2", 
                                   "Death Count"="#E7D366")) +
  facet_wrap(~region, scales = "free_y", shrink = TRUE) +
  scale_y_continuous(labels = comma) +
  theme(legend.title = element_blank(), plot.title.position = "middle")
  

ggplotly(fig2, tooltip = c("x", "y"))
```

```{r test, echo = FALSE}
mergedCA[is.na(mergedCA)] = 0
```

$$\\[.1in]$$

### [Fig 3: Vaccination by Company]{.ul}

```{r fig3, echo = FALSE, warnings = FALSE, message=FALSE}
mergedCA3 <-
  mergedCA %>%
  group_by(region,date) %>%
  summarise(county = county,
            region = region,
            ave_moderna = ave(moderna_perc),
            ave_pfizer = ave(pfizer_perc),
            ave_jj = ave(jj_perc))

mergedCA3[is.na(mergedCA3)] = 0


fig3 <- 
  mergedCA3 %>%
  ggplot(aes(group = region)) +
  geom_line(aes(x = date, y = ave_moderna, color = "Moderna")) +
  geom_line(aes(x = date, y = ave_pfizer, color = "Pfizer")) +
  geom_line(aes(x = date, y = ave_jj, color = "J&J")) +
  scale_colour_manual(values = c("Moderna"="#C78888", 
                                   "Pfizer"="#88C7C2", 
                                   "J&J"="#E7D366")) +
    labs(title = "Vaccination Efforts by Company", 
         x = "Date", y = "Percent of Total Doses Administered") +
    theme(legend.title = element_blank(), axis.title.y = element_text(hjust = -1.5)) +
    facet_wrap(~region)

ggplotly(fig3, tooltip = c("x", "y"))
```
